---
title: "Air Passengers Time Series Example"
author: "Moaeed Sajid"
date: "09/01/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(forecast)
data ("AirPassengers")
```

## EDA

Time Series Data can be modelled using the ARIMA model.  ARIMA stands for Auto Regressive Integrated Moving Average.

Auto Regression suggests that the current value is dependent on it's preceding values.  so if the current value is x(t) then the AR component is x(t-1)*a where a is the fitted co-efficient

These previous terms are referred to as lag values (first, second...) which we will see later on.

We will be using the Passengers dataset from the built in R forecasts library.  

Model validation will be done using the Ljung-Box test

```{r airpassengers}
class("AirPassengers")
start (AirPassengers)
end(AirPassengers)
frequency(AirPassengers)
sum(is.na(AirPassengers))
summary(AirPassengers)
AirPassengers
```

## Plot with Trend Line using lm


```{r echo=TRUE}
plot(AirPassengers)
abline(reg=lm(AirPassengers~time(AirPassengers)))
cycle(AirPassengers)
```

## Time Series Decompostition

The plot below shows the observed data, overall trend, seasonal variation and the the irregularity.  

Before building a time series model the data needs to be stationary.  That is that it needs to have a constant mean and variance.

```{r echo=TRUE}
tsdata <- ts(AirPassengers, frequency = 12)
ddata <- decompose(tsdata, type = "multiplicative")
plot(ddata)
#plot(ddata$trend/type/seasonal/random)
```


## Boxplot by Cycle

Shows much more passengers during Jun-Sep

```{r  echo=TRUE}
boxplot(AirPassengers~ cycle(AirPassengers), xlab = "Date")
```

## Identify PDQ

Use the auto.arima method in R to identify best values to use for ARIMA, thefirst set of brackets.  
 
 - AR(p) is the number of Auto Regressive Terms
 - I(d) the degree of differencing is how many non seasonal differences are required to achieve stationarity
 - MA(q) is the error of the model as a combination of previous error terms e~t~
 
Second set are about seasonality 

```{r echo=TRUE}
auto.arima(AirPassengers, ic = "aic", trace = TRUE)
myFit <- auto.arima(AirPassengers)
myFit
```     

## Checking the model

Plot the...

- Residuals

In order to test wether the data is stationary we can use the following 2 parameters

- Auto Correlation Function
Used to determine how past and future points are related in a time series by showing how correlated points are with each other, based on on how many time steps they are separated  

- Partial Auto Correlation Function
The degree of association between two variables whilst adjusting the effect ofone or more additional variables.

We can show our model to be stationary by the acf and pacf values not crossing the blue siginificance boundary lines.


```{r echo=TRUE}
library(tseries)
plot.ts(myFit$residuals, main = "Residuals")

acf(ts(myFit$residuals), main = "ACF Residuals")
pacf(ts(myFit$residuals), main = "PACF Residuals")
```     
## Predicting the model

h in forecast command will decide how far ahead to forecast


```{r echo=TRUE}
apForecast <- forecast(myFit, level = 95,h = 10*12)
plot(apForecast, main = "Forecast Predictions Going Forward")
```
  
  
## Testing the model using box-test

Box test below gives us p values which are relatively high showing the model to be accurate


```{r echo=TRUE}
Box.test(myFit$residuals, lag = 5, type = "Ljung-Box")
Box.test(myFit$residuals, lag = 10, type = "Ljung-Box")
Box.test(myFit$residuals, lag = 15, type = "Ljung-Box")

```